\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\ {\footnotesize Bachelorarbeit}

### 2023-03-20 Paper/paper.tex
intrinsics table

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ulem} % underline
\usepackage{float} % H float specifier
\usepackage{hyperref} % hyperlinks
% \usepackage{hyphenat}
\usepackage{bookmark}
\usepackage{systeme} % systems of equation
\usepackage[utf8]{inputenc} % necessary for proper unicode support
\usepackage{parcolumns} % multiple column parallel typesetting
\usepackage[english]{babel} % comment this for Angelsächsisch
\usepackage{graphicx} % including graphics
\usepackage{sourcecodepro} % best font
\usepackage{listings} % source code listings
\usepackage{biblatex} % better citation
% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry

@@

Standard vector data types have the format \texttt{uintnxm\_t} with lane width
$n$ in bits and and lane count $m$. Array types of the format
\texttt{uintnxmxc\_t}, $c\in\{1,2,3,4\}$ are also defined which are used in
operations requiring multiple parameters like \texttt{TBL} or pairwise
load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
after the operation name to indicate operation on a 128-bit register and end
with the lane data format. Multiplying eight pairs of 16-bit numbers \texttt{a,b} for
example can be done via the following:

@@

The compiler allocates vector registers for \texttt{a}, \texttt{b} and
\texttt{result} and assembles the intrinsic to \texttt{MUL Vr.8H, Va.8H,
Vb.8H}. Necessary load and stores for the result and parameters are also
handled automatically.

@@

uint8x16_t shl(uint8x16_t v, int n)
{
        // TODO check out vst
        uint64_t l0 = vgetq_lane_u64(v, 0);
        uint64_t l1 = vgetq_lane_u64(v, 1);
        l1 = l1 << n | (l0 >> (64 - n));
        l0 <<= n;
        uint8x16_t ret;
        ret = vsetq_lane_u64(l0, ret, 0);
        ret = vsetq_lane_u64(l1, ret, 1);
        return ret;
}

uint8x16_t shr(uint8x16_t v, int n)
{
        uint64_t l0 = vgetq_lane_u64(v, 0);
        uint64_t l1 = vgetq_lane_u64(v, 1);
        l0 = l0 >> n | (((l1 << (64 - n)) >> (64 - n)) << (64 - n));
        l1 >>= n;
        uint8x16_t ret;
        ret = vsetq_lane_u64(l0, ret, 0);
        ret = vsetq_lane_u64(l1, ret, 1);
        return ret;
}

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ulem} % underline
\usepackage{float} % H float specifier
\usepackage{hyperref} % hyperlinks
\usepackage{tabularx} % tabular with column specifier X to break lines
% \usepackage{hyphenat}
\usepackage{bookmark}
\usepackage{systeme} % systems of equation
\usepackage[utf8]{inputenc} % necessary for proper unicode support
\usepackage{parcolumns} % multiple column parallel typesetting
\usepackage[english]{babel} % comment this for Angelsächsisch
\usepackage{graphicx} % including graphics
\usepackage{sourcecodepro} % best font
\usepackage{listings} % source code listings
\usepackage{biblatex} % better citation
% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry

@@

Standard vector data types have the format \texttt{uintnxm\_t} with lane width
$n$ in bits and and lane count $m$. Array types of the format
\texttt{uintnxmxc\_t}, $c\in\{2,3,4\}$ are also defined which are used in
operations requiring multiple parameters like \texttt{TBL} or pairwise
load/stores. Intrinsics start with \texttt{v}, optionally contain a \texttt{q}
after the operation name to indicate operation on a 128-bit register, and end
with the lane data format. Multiplying eight pairs of 16-bit numbers
\texttt{a,b} for example can be done via the following:

@@

In this case, the compiler allocates vector registers for \texttt{a},
\texttt{b} and \texttt{result} and assembles the intrinsic to \texttt{MUL
Vr.8H, Va.8H, Vb.8H}. Necessary load and stores for the result and parameters
are also handled automatically. Of special interest to us are the following
intrinsics, each existing in different variants with different lane widths and
also array types: \\

\begin{table}[h!]
    \centering
    \footnotesize
    \begin{tabularx}{\textwidth}{ll|X}
        Intrinsic && Description \\
        \hline
        \texttt{uint8x16\_t} & \texttt{vreinterpretq\_u8\_u64(uint64x2\_t)} & Explicit casting \\
        \hline
        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
        \hline
        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
        \hline
        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
        \hline
        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
        \hline
        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
        \hline
        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
        \hline
        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
        \hline
        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
        \hline
        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
        \hline
        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
    \end{tabularx}
\end{table}

The implementations closely follow the aforementioned strategies;
instruction-level optimization is left to the compiler.

@@

uint8x16_t shl(uint8x16_t v, int n)
{
        uint64_t l[2];
        vst1q_u64(l, v);
        l[1] = l[1] << n | (l[0] >> (64 - n));
        l[1] <<= n;
        return vreinterpretq_u8_u64(vld1q_u64(l));
}

uint8x16_t shr(uint8x16_t v, int n)
{
        uint64_t l[2];
        vst1q_u64(l, v);
        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
        l[1] >>= n;
        return vreinterpretq_u8_u64(vld1q_u64(l));
}
